---
title: "R Notebook"
output: html_notebook
author: Vaibhavi Mukadam
---

```{r}
library(fpp2)
#install.packages("forecast")
library(forecast)
library(TTR)
library(lubridate)
library(reshape2)
library(tidyr)
library(tseries)
library(tidyverse)
library(timetk)
#install.packages("tidyverse")
#install.packages("timetk")
```
```{r}
og = read.csv('timeseriesdata.csv')
#og = GLD
tail(og)
og$datetime = as.POSIXct(og$date,format = "%Y-%m-%d %H")
head(og)
df = og
#df$date= as.Date(df$datetime)
df$Year <- strftime(df$datetime, "%Y")
df$Month <- strftime(df$datetime, "%m")
df$Day<- strftime(df$datetime, "%d")
df$Time <- strftime(df$datetime, "%H")
df$Week <- strftime(df$datetime, "%W")


#df = separate(df, "date", c("Year", "Month", "Day"), sep = "-")

```
Box Plot 
```{r}
boxplot(df$count~df$Month,xlab="Month", ylab="Count")
boxplot(df$count~df$Week,xlab="Week", ylab="Count")
boxplot(df$count~df$Time,xlab="Hour", ylab="Count")
boxplot(df$count~df$Day,xlab="Day", ylab="Count")
```
Initial data 
```{r}
summary(df)
library(timetk)
#plot_time_series()
#wm_ts = xts(df$count, start=c(2020,0), end=c(2021,1))
df %>% plot_time_series(df$datetime, df$count)
df %>% plot_time_series(df$datetime, log(df$count))

```

```{r}
library(zoo)

hourly_ts <- zoo(
  x         = df[["count"]],
  order.by  = df[["datetime"]],
  frequency = 24
)

start(hourly_ts)
end(hourly_ts)
```
```{r}


```

```{r}
library(xts)
hts <- xts(df$count, order.by = df$datetime)
Acf(hts)
plot(hts)
```

```{r}
ggAcf(hts)
acf(hts)
Pacf(hts)
tsdisplay(hts) # acf and pacf of the timeseries. 
```
```{r}
plot(hts)
#decomp <-decompose(as.ts(hts, fequency = 10)) 
```
 Decomposition
```{r}
decom <- decompose(hts)
decom$type
plot(decom)
sd = seasadj(decom)
plot(sd)
plot(dts)
lines(sd, col="Red")
d = stlf(hts)
```
```{r}
nfc = naive(hts,h=24) # forecasting h points 
plot(nfc) #plotting the forecast 
plot.ts(nfc$x,nfc$fitted,xy.labels = FALSE,xy.lines = FALSE, main = "Actual vs Fitted") #actual
```
```{r}
snfc = snaive(hts,h=24) # forecasting h points 
plot(snfc) #plotting the forecast 
plot.ts(snfc$x,snfc$fitted,xy.labels = FALSE,xy.lines = FALSE, main = "Actual vs Fitted") #actual
```
```{r}
f1=forecast(snfc)
f1
accuracy(snfc)
```

```{r}
#Residual Analysis - 
checkresiduals(nfc)
plot(nfc$residuals)
hist(nfc$residuals)
```

```{r}
f=forecast(nfc)
f
accuracy(nfc)
```
Moving Averages
```{r}
#par(mfrow=c(2,2))

ma3=ma(hts, 3)
ma6=ma(hts, 6)
ma9=ma(hts, 9) 
ma12=ma(hts, 12) 

ma24=ma(hts, 24) 

plot(ma3 , main = "order 3")
lines(hts, col='Red')
plot(ma6,  main = "order 6")
lines(hts, col='Blue')
plot(ma9,  main = "order 9")
lines(hts, col='Green')
plot(hts,  main = "order 12")
lines(ma12, col='Green')
library("scales")
plot(hts,  main = "order 24")
lines(ma24, col='Yellow', col = alpha(ma24, 0.5))

p <- ggplot(ma24, aes(x=ma24))
p + geom_line(aes(color=hts), stat="density", size=2, alpha=0.4)
```

```{r}

#forecast for 12 
ma3fc <- forecast(ma3,h=12)
plot(ma3fc)
accuracy(ma3fc)

ma6fc <- forecast(ma6,h=12)
plot(ma6fc)
accuracy(ma6fc)

#Analysis - 
checkresiduals(ma6fc)
Acf(ma6fc$residuals) #Check if there is random noise
dtslm = tslm(dts~trend+season)
plot.ts(ma6fc$x,ma6fc$fitted,xy.labels = FALSE,xy.lines = FALSE, main = "Actual vs Fitted") #actual
abline(0, 1, col="red")
accuracy(ma6fc)
```


```{r}
#tsdisplay(hts)
adf.test(hts)# ADF test says differences is required if p-value is > 0.05
#p-value = 0.621
kpss.test(hts)# Kipps test says differences is required if p-value is < 0.05
#p-value = 0.01
## Analysis: Differences is required

# NSDIFFS only works for seasonal data
#nsdiffs(hts)
# However NDIFFS works with non-seasonal data
ndiffs(hts)
```

```{r}
htsdiff1 <- diff(hts, differences=1) 
tsdisplay(htsdiff1)
Acf(htsdiff1) #Should look like white noise otherwise repeat 
ndiffs(htsdiff1)#p-value = 
kpss.test(htsdiff1)#p-value = 0.04221 it is less than 0.05 >>done
adf.test(htsdiff1) #p-value = 0.425 it is greated that 0.05
```
```{r}
#2nd
htsdiff2 <- diff(htsdiff1, differences=1) 
tsdisplay(htsdiff2)
Acf(htsdiff2) #Should look like white noise otherwise repeat 
ndiffs(htsdiff2)
kpss.test(htsdiff2) #p-value = 0.1 less than 0.05  >>done
adf.test(htsdiff2) # p-value = 0.02324 less than 0.05 >>done
#conclude d value - > 2
```

```{r}
#•	Plot the Time Series chart of the differences series. 
Acf(htsdiff1, lag.max=20) # lag 1
Acf(htsdiff1, lag.max=20,plot=FALSE)
Pacf(htsdiff1, lag.max=20) # lag 1
Pacf(htsdiff1, lag.max=20,plot=FALSE)
```

Acf(htsdiff2, lag.max=20) # lag at 2
Acf(htsdiff2, lag.max=20,plot=FALSE)
Pacf(htsdiff2, lag.max=20) # lag at 2
Pacf(htsdiff2, lag.max=20,plot=FALSE)

Time series with trends, or with seasonality, are not stationary — the trend and seasonality will affect the value of the time series at different times. 
#difference between acf and Acf is that Acf does not plot a spike at lag0.

Auto Arima 

```{r}
auto_fit = auto.arima(hts, trace = TRUE, stepwise = FALSE)
auto_fit # ARIMA(2,2,0)
#plot(auto_fit)
pacf(hts)
```

#Residual analsysis 
```{r}
arimaf =forecast(auto_fit)
arimaf

plot(arimaf)
plot.ts(arimaf$fitted,arimaf$x,xy.labels = FALSE,xy.lines = FALSE, main = "Residual vs Fitted") #actual
abline(0, 1, col="red")

accuracy(arimaf)
checkresiduals(arimaf)
```

```{r}

#Residual Analysis - 
checkresiduals(auto_fit)
plot(arimaf$residuals)
hist(arimaf$residuals)

Acf(arimaf$residuals) #Check if there is random noise
plot.ts(arimaf$residuals,arimaf$fitted,xy.labels = FALSE,xy.lines = FALSE, main = "Residual vs Fitted") #actual
abline(0, 1, col="red")

plot.ts(arimaf$residuals,arimaf$x,xy.labels = FALSE,xy.lines = FALSE, main = "Residual vs Fitted") #actual
abline(0, 1, col="red")
```